### 2022-09-25 Thesis/text/einleitung.tex
cleanup of tjhe doc

Core Problem - too little autonomy in a standard autonomous mobile robot running ROS navigation(AMR)

Description of the Problem

Robots regurlarly needs operator in case smth unexpected happens\\

this may include: unexpected collisions, a system restart is needed, battery is empty\\

The system is not able to react and handle unforeseen situations\\

This can lead to safety issues during the operation of a robot. These safety issues can lead to uncontrolled driving maneuvers due to wrong sensor data or a wrong environment represantation.\\

Due to the need of the robot to require an operator, the robot is closer to an automated instead of an autonomous vehicle. 
Difference is defined by the SAE Levels for automonous driving in cars.\\

It is important for the robot to possess the ability to navigate in an uncertain environment to guarantee the safety of the robot itself and all other actors in its environment, these can be other robots and humans.\\

In theory the robot can perform fully autonomous navigation including mapping and localization but as soon as something out of the ordinary happens the robots autonomy is not guaranteed to be reliable anymore.
Default robot is not able to represent all possible fail states and does not detect failures on a systematic level, but only inside its navigation related subsystem. And even when failures inside this subsystems are discovered, the options to handle these problems are very limited and often do not deal with the problems in an autonomous way. The reliance on human-needed problem solving decreases the robots usefulness when tasked with real goals.

Having a robust and safe robot behaviour opens the door to many applications in a more diverse set of challenging environments. Usually a reinforcement learning approach with broad training data set is a great way to handle a multitude of unknown environments and situations, but one of the biggest problems is the lacking determinism in such systems.

In the robotic world exists a need for a deterministic system which can enhance and override the decision making process and navigation capabilities of the robot.

This thesis proposes a comprehensive approach for behavior planning for an autonomous mobile robot.

\textbf{How can behavior planning increase the safety and autonomy of an autonomous mobile robot running ROS?}

>>>>

Mobile robots are becoming more widespread in many domains. They are marketed as autonomous mobile robots, but often the autonomy of the robot is limited to very specific environments and require frequent human help to function correctly. This hardly qualifies the robot to be labeled autonomous. 
One of the quickest and most popular ways to build and program a robot is to use the Robot Operating System (ROS). ROS offers many tools and functionalities to create and program robots, but even then a normal robot needs human supervision during the operation to ensure the proper function and safety. 
There are ROS packages to safely navigate from point A to point B, but currently they are lacking the ability to react to unforeseen events. It is important for the robot to possess the ability to navigate in an uncertain environment to guarantee the safety of the robot itself and all other actors in its environment, these can be other robots and humans. 
In theory the robot can perform fully autonomous navigation including mapping and localization but as soon as something out of the ordinary happens the robots autonomy is not guaranteed to be reliable anymore. 
Default robot is not able to represent all possible fail states and does not detect failures on a systematic level, but only inside its navigation related subsystem. And even when failures inside this subsystems are discovered, the options to handle these problems are very limited and often do not deal with the problems in an autonomous way. The reliance on human-needed problem solving decreases the robots usefulness when tasked with real goals. 
Therefore, this thesis centers around the research question how behavior planning can increase the robustness and autonomy of a robot running ROS2. 
Behavior planning enables the robot to react to failures and problems of the system and can decide alternative courses of action to mitigate risks and finish the given tasks.

%Core Problem - too little autonomy in a standard autonomous mobile robot running ROS navigation(AMR)
%
%Description of the Problem:
%
%Despite the advancements in autonomous driving the robot regurlarly needs an operator in case smth unexpected happens.
%
%This may include: unexpected collisions, a system restart is needed, battery is empty, the environment is too crowded and the robot gets stuck
%
%The system is not able to react and handle unforeseen situations.
%
%This can lead to safety issues during the operation of a robot. These safety issues can lead to uncontrolled driving maneuvers due to wrong sensor data or a wrong environment represantation.
%
%Due to the need of the robot to require an operator for safe, the robot is not really autonomous despite it being labeled as such. 
%
%(Difference is defined by the SAE Levels for automonous driving in cars.) \\
%
%Relevance, need for research:\\
%\\

%
%Having a robust and safe robot behaviour opens the door to many applications in a more diverse set of challenging environments. Usually a reinforcement learning approach with broad training data set is a great way to handle a multitude of unknown environments and situations, but one of the biggest problems is the lacking determinism in such systems.
%
%In the robotic world exists a need for a deterministic system which can enhance and override the decision making process and navigation capabilities of the robot. 
%
%Planned solution:\\

This thesis will explore possibilities how current systems can be improved through the addition of a dedicated behavior planning component. Furthermore, the thesis will provide an implementation of examplatory behaviors and provide a system architecture to incorporate behavior planning in other robots, too

%The thesis will explore how an extension of the current software architecture and behavior planning capabilities for an autonomous mobile robot that is running ROS2 can improve the autonomy of the robot. 
%
%Expected results:\\

The desired outcomes of these implemented software will be an increase of robustness and decrease of required human actions during a number of scenarios. In these scenarios, failures and problems are artificially induced to test the robots abilities to behave autonomously.

%\textit{Research gap is not named so far}

### 2022-09-25 Thesis/text/ergebnisse.tex
cleanup of tjhe doc

To test the efficacy of the behavior planning approach in relation to the requirements from chapter 3.xx, the robot is put into specific scenarios that should trigger the behavior responses.

The scenarios are executed in a simulated apartment environment as seen in figure xx and in a simulated world with small round obstacles in an enclosed space as depicted in figure xx.

\begin{figure}
	\includegraphics[width=0.5\textwidth]{images/house_env.png}
	\caption{Apartment Environment in Gazebo}
\end{figure}

\begin{figure}
	\includegraphics[width=0.5\textwidth]{images/world_env.png}
	\caption{Enclosed Environment in Gazebo}
\end{figure}

@@

The scenarios are derived from the functional requirements and can be used to test multiple requirements.
The acceptance criteria often contains multiple binary conditions that can only be met or failed. All acceptance criteria for a given scenario must be met, for the test to be counted as success.

@@

\begin{table}[h!]
\caption{Scenarios}
	\begin{tabular}{| m{0.1\textwidth} | m{0.13\textwidth}| m{0.33\textwidth} | m{0.34\textwidth}|} 
  	\hline
  	Name & Related Requirements & Description & Success Criteria\\ 
  	\hline
  	Lidar$\_$1 & fn$\_$req1, fn$\_$req2, fn$\_$req3, fn$\_$req4, fn$\_$req5 & Robot is standing still, Lidar node crashes & Reset the system\\ 
  	\hline
  	Lidar$\_$2 & fn$\_$req1, fn$\_$req2, fn$\_$req3, fn$\_$req4, fn$\_$req5 & Navigating with 0.25 m/s straight towards goal (1m away) & Reset system, reach goal \\ 
  	\hline
  	Lidar$\_$3 & fn$\_$req1, fn$\_$req2, fn$\_$req3, fn$\_$req4, fn$\_$req5 & Navigating with 0.25 m/s and 0.5 rad/s towards goal (1m away) & Reset system, reach goal\\
  	\hline
  	IMU$\_$1 & fn$\_$req1, fn$\_$req2, fn$\_$req3, fn$\_$req4, fn$\_$req5 & Robot is standing still, IMU node crashes & Reset system \\
  	\hline
  	IMU$\_$2 & fn$\_$req1, fn$\_$req2, fn$\_$req3, fn$\_$req4, fn$\_$req5 & Navigating with 0.25 m/s straight towards goal (1m away) & Reset system, reach goal \\ 
  	\hline
  	IMU$\_$3 & fn$\_$req1, fn$\_$req2, fn$\_$req3, fn$\_$req4, fn$\_$req5 & Navigating with 0.25 m/s and 0.5 rad/s towards goal (1m away) & Reset system, reach goal\\
  	\hline
  	Odom$\_$1 & fn$\_$req1, fn$\_$req2, fn$\_$req3, fn$\_$req4, fn$\_$req5 & Robot is standing still, Odom node crashes & Reset the system\\ 
  	\hline
  	Odom$\_$2 & fn$\_$req1, fn$\_$req2, fn$\_$req3, fn$\_$req4, fn$\_$req5 & Navigating with 0.25 m/s straight towards goal (1m away) & Reset system, reach goal \\ 
  	\hline
  	Odom$\_$3 & fn$\_$req1, fn$\_$req2, fn$\_$req3, fn$\_$req4, fn$\_$req5 & Navigating with 0.25 m/s and 0.5 rad/s towards goal (1m away) & Reset system, reach goal\\
  	\hline
 	\end{tabular}
\end{table}

\begin{table}[h!]
	\caption{Scenarios}
	\begin{tabular}{| m{0.1\textwidth} | m{0.13\textwidth}| m{0.33\textwidth} | m{0.34\textwidth}|} 
  	\hline	
  	NoPath Found$\_$1 & fn$\_$req7 & Path to goal can not be calculated, robot is standing still &  
Reset system, calculate path to goal, reach goal \\
	\hline
	NoPath Found$\_$2 & fn$\_$req7, fn$\_$req8 & Goal is unreachable, robot is standing still & Alternative goals, close to the original are tested to be reachable \\
	\hline
	Collision$\_$1 & fn$\_$req2, fn$\_$req4, fn$\_$req6 & Robot is standing still and a collision with the robot is caused & Robot can get out of collision state, navigation to goals still working \\
	\hline
	Collision$\_$2 & fn$\_$req2, fn$\_$req4, fn$\_$req6 & The robot is driving and collides with an undetected obstacle (0.25m/s) &  
Robot can get out of collision state, navigation to goals still working, undetected obstacle gets added to map \\
	\hline
	Battery$\_$1 & fn$\_$req9 & The robot battery runs low & The robot will not drive to a goal which is outside of its reachable range \\
	\hline
	Motor$\_$1 & fn$\_$req2, fn$\_$req5 & Hardware failure on the motors & n.a. \\
	\hline

@@

The results in table xx show an increase in the successful handling of the test scenarios for most test cases. The system supervision component is a very reliable component and enables the system to recover from an unexpected sensor failure. The more advanced behaviors that were implemented also show good improvements in the autonomous handling of the scenarios. The Battery Scenario was successful every time it was induced. However, the path planning behavior for finding alternative goals was not successful in about 50 percent of the time. The carried out test showed that alternative methods for finding new goals would probably have resulted in better performance. The collision behavior during driving worked really well and the map updates are a important feature for the success of the scenarios. Nevertheless, there is a lack of intelligence and autonomy when the robot is standing still and a collision is forced as it relies on the same mechanism as during the driving scenario. This led to a case in which the robot reversed into a wall because the collision turned the robot by about 90 degrees. This led to a failed test in which a operator would be needed to correct the robot. This apparent lack of robustness was not apparent in the driving collision scenario because the force of the robot was not enough to cause a major shift in orientation.

\begin{table}[h!]
\caption{Results}
	\begin{tabular}{| m{0.11\textwidth} | m{0.13\textwidth}| m{0.33\textwidth} | m{0.33\textwidth}|} 
  	\hline
  	Name & Number of Runs &  Percentage Succesful Without Behavior Planning & Percentage Succesful With Behavior Planning\\ 
  	\hline
  	Lidar$\_$1 & 5 & 0 & 100 \\ 
  	\hline
  	Lidar$\_$2 & 5 & 0 & 100 \\ 
  	\hline
  	Lidar$\_$3 & 5 & 0 & 100 \\ 
  	\hline
  	IMU$\_$1 & 5 & 0 & 100 \\ 
  	\hline
  	IMU$\_$2 & 5 & 0 & 100 \\ 
  	\hline
  	IMU$\_$3 & 5 & 0 & 100 \\ 
  	\hline
  	Odom$\_$1 & 5 & 0 & 100 \\ 
  	\hline
  	Odom$\_$2 & 5 & 0 & 100 \\ 
  	\hline
  	Odom$\_$3 & 5 & 0 & 100 \\ 
  	\hline
  	NoPath Found$\_$1 & 20 & 0 & 100 \\ 
  	\hline
  	NoPath Found$\_$2 & 20 & 0 & 55 \\ 
  	\hline
  	Collision$\_$1 & 20 & 0 & 95 \\ 
  	\hline
  	Collision$\_$2 & 20 & 0 & 100 \\ 
  	\hline
  	Battery$\_$1 & 10 & 0 & 100 \\ 
  	\hline
  	Motor$\_$1 & 10 & 0 & 0 \\ 
  	\hline
	\end{tabular}
\end{table}

The non-functional requirements are mostly met by the architectural design choices. 
The implementation of the whole behavior planning and system supervision system is eliminating a single point of failure for the system. Even when the whole Navigation2 system collapses is the robot able to move to a degree and more importantly stop completely (compare table 3.3, non$\_$fn$\_$req). The behavior tree itself and algorithms that are implemented in the behavior tree are all deterministic (non$\_$fn$\_$req3). The robots behavior goes beyond pure reactivity when navigating through environments, as the advanced behaviors are making use of a dedicated planning phase before the actions are carried out (non$\_$fn$\_$req4). 
And finally, the performance of the system when checking the conditions of the BT is below the desired threshold of 10ms, as the whole system runs in multiple threads which allows the behavior tree to complete one complete cycle in about 5ms with current system (non$\_$fn$\_$req2).

>>>>

To test the efficacy of the behavior planning approach in relation to the requirements from chapter 3.2, the robot is put into specific scenarios that should trigger the behavior responses.

The scenarios are executed in a simulated apartment environment as seen in figure \ref{fig:house_gazebo} and in a simulated world with small round obstacles in an enclosed space as depicted in figure \ref{fig:world_gazebo}. 
\begin{center}
\begin{figure}[ht]
	\label{fig:house_gazebo}
	\includegraphics[width=0.75\textwidth]{images/house_env.png}
	\caption{Apartment Environment in Gazebo}
\end{figure}
\end{center}

\begin{center}
\begin{figure}
	\label{fig:world_gazebo}
	\includegraphics[width=0.75\textwidth]{images/world_env.png}
	\caption{Enclosed Environment in Gazebo}
\end{figure}
\end{center}

@@

The scenarios are derived from the functional requirements and can be used to test multiple requirements and listed in the tables \ref{tab:sensor_scenarios} and \ref{tab:behavior_scenarios}.
The acceptance criteria often contains multiple binary conditions that can only be met or failed. All acceptance criteria for a given scenario must be met, for the test to be counted as success.

@@

\begin{table}[ht]
\label{tab:sensor_scenarios}
\caption{Scenarios}
	\begin{tabular}{| m{0.1\textwidth} | m{0.13\textwidth}| m{0.33\textwidth} | m{0.34\textwidth}|} 
  	\hline
  	Name & Related Requirements & Description & Success Criteria\\ 
  	\hline
  	Lidar$\_$1 & fn$\_$req1, fn$\_$req2, fn$\_$req3, fn$\_$req4, fn$\_$req5 & Robot is standing still, Lidar node crashes & Reset the system\\ 
  	\hline
  	Lidar$\_$2 & fn$\_$req1, fn$\_$req2, fn$\_$req3, fn$\_$req4, fn$\_$req5 & Navigating with 0.25 m/s straight towards goal (1m away) & Reset system, reach goal \\ 
  	\hline
  	Lidar$\_$3 & fn$\_$req1, fn$\_$req2, fn$\_$req3, fn$\_$req4, fn$\_$req5 & Navigating with 0.25 m/s and 0.5 rad/s towards goal (1m away) & Reset system, reach goal\\
  	\hline
  	IMU$\_$1 & fn$\_$req1, fn$\_$req2, fn$\_$req3, fn$\_$req4, fn$\_$req5 & Robot is standing still, IMU node crashes & Reset system \\
  	\hline
  	IMU$\_$2 & fn$\_$req1, fn$\_$req2, fn$\_$req3, fn$\_$req4, fn$\_$req5 & Navigating with 0.25 m/s straight towards goal (1m away) & Reset system, reach goal \\ 
  	\hline
  	IMU$\_$3 & fn$\_$req1, fn$\_$req2, fn$\_$req3, fn$\_$req4, fn$\_$req5 & Navigating with 0.25 m/s and 0.5 rad/s towards goal (1m away) & Reset system, reach goal\\
  	\hline
  	Odom$\_$1 & fn$\_$req1, fn$\_$req2, fn$\_$req3, fn$\_$req4, fn$\_$req5 & Robot is standing still, Odom node crashes & Reset the system\\ 
  	\hline
  	Odom$\_$2 & fn$\_$req1, fn$\_$req2, fn$\_$req3, fn$\_$req4, fn$\_$req5 & Navigating with 0.25 m/s straight towards goal (1m away) & Reset system, reach goal \\ 
  	\hline
  	Odom$\_$3 & fn$\_$req1, fn$\_$req2, fn$\_$req3, fn$\_$req4, fn$\_$req5 & Navigating with 0.25 m/s and 0.5 rad/s towards goal (1m away) & Reset system, reach goal\\
  	\hline
 	\end{tabular}
\end{table}

\begin{table}[ht]
	\label{tab:behavior_scenarios}
	\caption{Scenarios}
	\begin{tabular}{| m{0.1\textwidth} | m{0.13\textwidth}| m{0.33\textwidth} | m{0.34\textwidth}|} 
	\hline
	Name & Related Requirements & Description & Success Criteria\\ 
  	\hline	
  	NoPath Found$\_$1 & fn$\_$req7 & Path to goal can not be calculated, robot is standing still &  
Reset system, calculate path to goal, reach goal \\
	\hline
	NoPath Found$\_$2 & fn$\_$req7, fn$\_$req8 & Goal is unreachable, robot is standing still & Alternative goals, close to the original are tested to be reachable \\
	\hline
	Collision$\_$1 & fn$\_$req2, fn$\_$req4, fn$\_$req6 & Robot is standing still and a collision with the robot is caused & Robot can get out of collision state, navigation to goals still working \\
	\hline
	Collision$\_$2 & fn$\_$req2, fn$\_$req4, fn$\_$req6 & The robot is driving and collides with an undetected obstacle (0.25m/s) &  
Robot can get out of collision state, navigation to goals still working, undetected obstacle gets added to map \\
	\hline
	Battery$\_$1 & fn$\_$req9 & The robot battery runs low & The robot will not drive to a goal which is outside of its reachable range \\
	\hline
	Motor$\_$1 & fn$\_$req2, fn$\_$req5 & Hardware failure on the motors & n.a. \\
	\hline

@@

The results in table \ref{tab:results} show an increase in the successful handling of the test scenarios for most test cases. The system supervision component is a very reliable component and enables the system to recover from an unexpected sensor failure. The more advanced behaviors that were implemented also show good improvements in the autonomous handling of the scenarios. The Battery Scenario was successful every time it was induced. However, the path planning behavior for finding alternative goals was not successful in about 50 percent of the time. The carried out test showed that alternative methods for finding new goals would probably have resulted in better performance. The collision behavior during driving worked really well and the map updates are a important feature for the success of the scenarios. Nevertheless, there is a lack of intelligence and autonomy when the robot is standing still and a collision is forced as it relies on the same mechanism as during the driving scenario. This led to a case in which the robot reversed into a wall because the collision turned the robot by about 90 degrees. This led to a failed test in which a operator would be needed to correct the robot. This apparent lack of robustness was not apparent in the driving collision scenario because the force of the robot was not enough to cause a major shift in orientation.

\begin{table}[ht]
\label{tab:results}
\caption{Results}
	\begin{tabular}{| m{0.11\textwidth} | m{0.13\textwidth}| m{0.33\textwidth} | m{0.33\textwidth}|} 
  	\hline
  	Name & Number of Runs &  Percentage Succesful Without Behavior Planning & Percentage Succesful With Behavior Planning\\ 
  	\hline
  	Lidar$\_$1 & 5 & 0 & 100 \\ 
  	\hline
  	Lidar$\_$2 & 5 & 0 & 100 \\ 
  	\hline
  	Lidar$\_$3 & 5 & 0 & 100 \\ 
  	\hline
  	IMU$\_$1 & 5 & 0 & 100 \\ 
  	\hline
  	IMU$\_$2 & 5 & 0 & 100 \\ 
  	\hline
  	IMU$\_$3 & 5 & 0 & 100 \\ 
  	\hline
  	Odom$\_$1 & 5 & 0 & 100 \\ 
  	\hline
  	Odom$\_$2 & 5 & 0 & 100 \\ 
  	\hline
  	Odom$\_$3 & 5 & 0 & 100 \\ 
  	\hline
  	NoPath Found$\_$1 & 20 & 0 & 100 \\ 
  	\hline
  	NoPath Found$\_$2 & 20 & 0 & 55 \\ 
  	\hline
  	Collision$\_$1 & 20 & 0 & 95 \\ 
  	\hline
  	Collision$\_$2 & 20 & 0 & 100 \\ 
  	\hline
  	Battery$\_$1 & 10 & 0 & 100 \\ 
  	\hline
  	Motor$\_$1 & 10 & 0 & 0 \\ 
  	\hline
	\end{tabular}
\end{table}

The non-functional requirements are mostly met by the architectural design choices. 
The implementation of the whole behavior planning and system supervision system is eliminating a single point of failure for the system. Even when the whole Navigation2 system collapses is the robot able to move to a degree and more importantly stop completely (compare table \ref{tab:tab:nonfn_req}, non$\_$fn$\_$req). The behavior tree itself and algorithms that are implemented in the behavior tree are all deterministic (non$\_$fn$\_$req3). The robots behavior goes beyond pure reactivity when navigating through environments, as the advanced behaviors are making use of a dedicated planning phase before the actions are carried out (non$\_$fn$\_$req4). 
And finally, the performance of the system when checking the conditions of the BT is below the desired threshold of 10ms, as the whole system runs in multiple threads which allows the behavior tree to complete one complete cycle in about 5ms with current system (non$\_$fn$\_$req2).
